---
title: "vegan examples"
output: html_document
---

## 1. Libraries
```{r Libraries, echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
library(vegan)
library(tidyverse)
library(ggvegan)
```



## 2. Load data (mite examples from package in this instance)
```{r, echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
data(mite)
data(mite.env)
data(mite.pcnm)
```


## 3. Unconstrained ordination example

### Principal component analysis
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
pca <- rda(mite.env[,1:2]) #solo las primeras dos columnas de este data frame
summary(pca)
autoplot(pca, arrows = TRUE)
```


## 4. Constrained ordination example
### Redundancy analysis
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
rda <- rda(mite~SubsDens+WatrCont, data=mite.env)
rda
autoplot(rda, arrows = TRUE)
```

#### You can use more than one predictor variable matrix
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
rda2 <- rda(mite~.+as.matrix(mite.pcnm[,1:3]),data=mite.env)
rda2
autoplot(rda2, arrows = TRUE)
```

#### You can also partition the variation among space and environment
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
v.part <- varpart(mite, mite.env[,1:2], mite.pcnm[,1:3])
v.part
plot(v.part)
```


#### We can try with variables divided into more groups
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
v.part2 <- varpart(mite, ~ SubsDens + WatrCont, ~ Substrate + Shrub + Topo, mite.pcnm[,1:3], data = mite.env, transfo = "hel")
v.part2
plot(v.part2)
```


## 5. Non-metric Multidimensional Scaling
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
nmds <- metaMDS(mite)
autoplot(nmds, geom = "text", legend = "none")
```


#### Alternative plot
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
col = c("red", "blue", "green")
shape = c(18,20,15)
plot(nmds$points, col = col[mite.env$Shrub], pch = shape[mite.env$Shrub], cex = 1.2, main = "Mite Shrub Groups", xlab = "NMDS 1", ylab = "NMDS 2")
ordispider(nmds, groups = mite.env$Shrub, label = TRUE)
```

### You can test for differences in communities among the shrub groups
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
test1 <- adonis(mite~Shrub, data = mite.env, permutations = 999, method = "bray")
test1
```


### Crtitical asssumption for adonis is that variance is equal among the groups tested for a condition (Shrub in this case). We can test for this
```{r , echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
distances.mite <- vegdist(mite)
test2 <- anova(betadisper(distances.mite, mite.env$Shrub))
test2
test3 <- envfit(nmds~SubsDens + WatrCont, mite.env)
test3
plot(test3)
```


























